---
editor_options:
  markdown:
    wrap: 72
---

------------------------------------------------------------------------

---
---
title: "analysis.R"
author: "Saie"
date: "2023-07-05"
output:
  html_document:
    theme: 
---

### Installing packages

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### Importing dataset

```{r}
data <- read_csv("dailyActivity_merged.csv")
head(data)  # View the first few rows of the dataset
summary(data)  # View summary statistics of the variables
```

### Data Cleaning and Processing

#### Impute missing values with mean

```{r}
data$TotalSteps[is.na(data$TotalSteps)] <- mean(data$TotalSteps, na.rm = TRUE)
```

#### Converting to proper format

```{r}
data$ActivityDate <- as.Date(data$ActivityDate, format = "%m/%d/%Y")
```

```{r}
library(dplyr)
```

#### Renaming Variables

```{r}
dailyActivity_merged <- data %>%
  rename(Total_Distance = TotalDistance)
```

#### Removing duplicates

```{r}
dailyActivity_merged <- distinct(dailyActivity_merged)
```

# Data transformation

```{r}
dailyActivity_merged$log_TotalSteps <- log(dailyActivity_merged$TotalSteps)
```

#### Creating derived variables

```{r}
daily_steps <- dailyActivity_merged %>%
  group_by(ActivityDate) %>%
  summarise(Avg_Steps = mean(TotalSteps))
```

#### Data aggregation

```{r}
weekly_averages <- dailyActivity_merged %>%
  group_by(week = lubridate::week(ActivityDate)) %>%
  summarise(Avg_Steps = mean(TotalSteps))
```

## Exploring the variables

```{r}
# Histogram 
library(ggplot2)
ggplot(dailyActivity_merged, aes(x = TotalSteps)) +
  geom_histogram()
```

```{r}
# Line Chart 
ggplot(dailyActivity_merged, aes(x = ActivityDate, y = Total_Distance, color = ActivityDate, fill = Total_Distance)) +
  geom_line()
```

```{r}
# Scatter Plots
ggplot(dailyActivity_merged, aes(x = Total_Distance, y = Calories, color = Calories, alpha = Total_Distance)) +
  geom_point()
```

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
# Convert ActivityDate column to Date format
dailyActivity_merged$ActivityDate <- as.Date(dailyActivity_merged$ActivityDate)
```

## Time-based Trends

```{r}
ggplot(dailyActivity_merged, aes(x = ActivityDate, y = TotalSteps, fill = ActivityDate)) +
  geom_line() +
  labs(x = "Date", y = "Total Steps", color = "red") +
  ggtitle("Trends in Total Steps")
```

```{r}
ggplot(dailyActivity_merged, aes(x = ActivityDate, y = Total_Distance)) +
  geom_line() +
  labs(x = "Date", y = "Total Distance", color = 'blue') +
  ggtitle("Trends in Total Distance")
```

```{r}
ggplot(dailyActivity_merged, aes(x = ActivityDate, y = Calories)) +
  geom_line() +
  labs(x = "Date", y = "Calories") +
  ggtitle("Trends in Calories")
```

## Seasonal Patterns

```{r}
dailyActivity_merged$Month <- month(dailyActivity_merged$ActivityDate, label = TRUE)
monthly_avg_steps <- dailyActivity_merged %>%
  group_by(Month) %>%
  summarize(avg_steps = mean(TotalSteps))
```

```{r}
ggplot(monthly_avg_steps, aes(x = Month, y = avg_steps)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(x = "Month", y = "Average Steps") +
  ggtitle("Average Steps by Month")
```

# Correlations

```{r}
correlation <- cor(dailyActivity_merged$TotalSteps, dailyActivity_merged$Calories)
print(correlation)
```

# Long-term Trends

```{r}
weekly_avg_steps <- dailyActivity_merged %>%
  group_by(week = floor_date(ActivityDate, "week")) %>%
  summarize(avg_steps = mean(TotalSteps))
```

```{r}
ggplot(weekly_avg_steps, aes(x = week, y = avg_steps, color = week)) +
  geom_line() +
  labs(x = "Week", y = "Average Steps") +
  ggtitle("Trends in Average Steps (Weekly)")
```

## Additional Analysis

```{r}
activity_levels <- dailyActivity_merged %>%
  summarise(sedentary = sum(SedentaryMinutes),
            light_activity = sum(LightlyActiveMinutes),
            moderate_activity = sum(FairlyActiveMinutes),
            vigorous_activity = sum(VeryActiveMinutes))
```

### Percentage of time spent in each activity level

```{r}
total_minutes <- sum(activity_levels)
activity_levels_pct <- activity_levels / total_minutes * 100
print(activity_levels_pct)
```

## Sedentary: 81.33%

##Light Activity: 15.82% ##Moderate Activity: 1.11% ##Vigorous Activity:
1.74% ##These percentages indicate the distribution of activity levels
among the users in the dataset. It shows that a large portion of the
recorded time is spent in sedentary activities, while a smaller
percentage is allocated to light, moderate, and vigorous activities
